The abundance of Bullet-Groups in $\Lambda$CDM

### 2014-03-11 paper/ms.tex
subsection 4.1

The Bullet Cluster (massive cluster of galaxies \bulla) provided a new
kind of observational evidence for the existence of dark matter
\citep{Markevitch2004,Clowe2006}. Quantifying the displacement between
dark matter and the dominant baryonic component (hot X-ray emitting gas)
has been used to test the Cold Dark Matter (CDM)
paradigm itself by quantifying the substructure velocity required to
produce such displacement \citep{Hayashi2006, Springel2007,
  Thompson2012}, the displacemeent between the dominant dark matter
and baryonic component \citep{ForeroRomero2010} to estimate the
expected abundance of such events in a $\Lambda$CDM Universe and even
explore possible extensions to the concordance cosmological model
\citep{Farrar2007,Lee2010,Lee2012}.

@@

The Bullet-like configurations are composed by two dark matter
structures: the host halo and the dominant sub-halo. We describe the
kinematics of this configuration by the 
position and velocity vectors of the sub-halo in a frame of reference
where the main halo is at rest; thus
$\vec{v{}}=\vec{v}_{sub}-\vec{v}_{halo}$ and
$\vec{r}=\vec{r}_{sub}-\vec{r}_{halo}$, where the subscripts $host$
and $sub$ refer to the host and sub-halo in the frame of reference of
the simulation, respectively.

@@

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement
  between the host halo and its largest sub-halo. The left panel
  shows the results in terms of the physical displacements while
  the right panel shows the displacements normalized by the virial radius of
  the host halo. The continuous line corresponds to the halos in the
  group sample $V_{\rm circ,host}>700\kms$ and the dashed lines to the
  sample $V_{\rm circ,host}<700\kms$.
  The vertical lines corresponds to mean estimate separation 
  between the two dark matter clumps in the results reported by
  \citet{Gastaldello} for the group SL2S J08544-0121. Between $50\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark.}
\label{fig:displacement}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps} 
\end{center}
\caption{2D histogram in the plane $V_{\rm c,sub}/V_{\rm
    c,host}$-$d_{2D}$. The left panel corresponds to clusters and the
  right panel to groups. The circle with error bars corresponds to
  \bullg data reported from \citet{2013A&A...552A..80M} and
  \citet{Gastaldello}. The data used to construct the histograms
  integrates the objects at all redshifts. }
\label{fig:mass_displacement}
\end{figure*}

@@

The panel with the physical displacements also shows a vertical stripe
with the estimated displacement for the Bullet-group reported by
\cite{Gastaldello}. Considering this system as consistent withe the
groups sample we see that a fraction of $\sim 15\pm 2\%$ of the groups
should present a displacement equal or larger than the one estimated
for \bullg, this fraction is close to $45\pm2\% $. The panel with the
normalized displacements shows a  distribution that can be considered
close to universal in the sense that the two samples (groups and
clusters) at all redshifts present a similar trend.[...]

Figure \ref{fig:mass_displacement} shows 2D histograms in a plane
composed by the ratio of the two circular velocities $V_{\rm c,
  sub}/V_{\rm c, host}$ and the physical displacements. This is
revealing because these two quantities can be constrained by
observations. The displacement is a direct observable, while the ratio
of the circular velocities can be estimated from lensing studies or
approximated by the ratio of the total galaxy luminosity associated
with the galaxy peaks.

To construct this figure we co-add all the halos in the sample (left,
groups; right, clusters) at all redshifts. We stack the data because
we do not observe any strong time evolution, additionally this allows
us to increase the signal in each bin. Overplotted there is an circle
with error bars that represents the observational estimates for the
system \bullg using the fraction in velocity dispersion in
the line-of-sigth of the group SL2S SJ08544-0121
($\sigma_{host}=341^{+43}_{-109}$\kms and
$\sigma_{sub}=185^{+30}_{-62}$\kms) reported by
\citet{2013A&A...552A..80M} and the DM displacement inferred from the
data presented by \cite{Gastaldello}.

>>>>

The Bullet Cluster (\bulla) provided a new
kind of observational evidence for the existence of dark matter
\citep{Markevitch2004,Clowe2006}. Quantifying the displacement between
dark matter and the dominant baryonic component (hot X-ray emitting gas)
has been used to test the Cold Dark Matter (CDM)
paradigm itself by quantifying the substructure velocity required to
produce such displacement \citep{Hayashi2006, Springel2007,
  Thompson2012}, the displacemeent between the dominant dark matter
and baryonic component \citep{ForeroRomero2010} to estimate the
expected abundance of such events in a $\Lambda$CDM Universe and even
explore possible extensions to the concordance cosmological model
\citep{Farrar2007,Lee2010,Lee2012}.

@@

The Bullet-like configurations are composed by two dark matter
structures: the host halo and the dominant sub-halo. We describe the
kinematics of this configuration by the 
position and velocity vectors of the sub-halo in a frame of reference
where the main halo is at rest; thus
$\vec{v{}}=\vec{v}_{sub}-\vec{v}_{halo}$ and
$\vec{r}=\vec{r}_{sub}-\vec{r}_{halo}$, where the subscripts $host$
and $sub$ refer to the position of the minium of potential for the
host and sub-halo in the frame of reference of the simulation,
respectively.

@@

The geometrical configuration can be further described by the
following quantities. The circular velocity of each component, $V_{\rm
  c,host}$ for the host and $V_{\rm c, sub}$ for the sub-halo and the
size of the host halo $R_{\rm vir}$. Another useful quantity computed
in the simulation is the distance between the minium of potential for
the host halo and its center of mass, $X_{\rm
  off}=||\vec{r}_{min}-\vec{r}_{cm}||/R_{\rm vir}$ which serves as a
measurement of how perturbed is the host halo.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.9\textwidth]{figure_1.eps}
\end{center}
\caption{
  Integrated probability distribution for the displacement
  between the center of the host halo and its dominant sub-halo. The left panel
  shows the results in terms of the physical displacements while
  the right panel shows the displacements normalized by the virial radius of
  the host halo. The continuous (dashed) line corresponds to the halos in the
  cluster (group).
  The vertical lines shows the mean and uncertainties in the estimated
  separation  between the two dark matter clumps in the results reported by
  \citet{Gastaldello} for the \bullg. Between $50\%$
  to $60\%$ of the groups show a displacement equal or larger than
  this observational benchmark.}
\label{fig:displacement}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.0\textwidth]{figure_2.eps} 
\end{center}
\caption{2D histogram in the plane $V_{\rm c,sub}/V_{\rm
    c,host}$-$d_{2D}$. The left panel corresponds to clusters and the
  right panel to groups. The star with error bars corresponds to
  \bullg data reported from \citet{2013A&A...552A..80M} and
  \citet{Gastaldello}. The data used to construct the histograms
  integrates the objects at all redshifts. }
\label{fig:mass_displacement}
\end{figure*}

@@

The panel with the projected 2D physical displacements also shows a
vertical stripe with the estimated displacement for the Bullet-group reported by
\cite{Gastaldello}. In the group sample we see that a fraction of
$50\%$ to $60\%$  should present a displacement equal than the
estimate for \bullg; in the cluster sample this fraction increases to
$80\%$-$85\%$. This fraction is naturally higher in more massive
systems because they are larger in size. Normalizing the displacements
by the virial radius, right panel Figure \ref{fig:displacement}, we
see that the distribution is the same regardless of the sample and the
redshift.

Figure \ref{fig:mass_displacement} shows 2D histograms in a plane
defined by the ratio of the two circular velocities $V_{\rm c,
  sub}/V_{\rm c, host}$ and the projected 2D physical displacements;
quantities that can be constrained by observations. The displacement
is a direct observable, while the ratio of the circular velocities can
be estimated from lensing studies or approximated by the ratio of the
total galaxy luminosity associated with the galaxy peaks or by a
lensing reconstruction.

To construct this figure we co-add all the halos in the sample (left,
clusters; right, groups) at all redshifts. We stack the data because
we do not observe any strong redshift dependence, additionally this allows
us to increase the signal in each bin. We overplot a star with error
bars that represents the observational estimates for the system
\bullg\ using the fraction in velocity dispersion in the line-of-sigth
of the group SL2S SJ08544-0121 ($\sigma_{host}=341^{+43}_{-109}$\kms and
$\sigma_{sub}=185^{+30}_{-62}$\kms) reported by
\citet{2013A&A...552A..80M} and the DM displacement inferred from the
data presented by \cite{Gastaldello}.
